The Name of the Title is Hope

### 2023-12-13 input/5_user_study_and_methods.tex
revision

\section{User study \& Methods}
With this study, we wanted to find out whether the experience of agency can be preserved during physical action augmentation when the user's own brain signals serve as the control signal to a physical end effector. To this end, we assessed whether our prototype preserved agency using a mixed methods approach including a psychometric test of intentional binding, one standardized question, and a qualitative (phenomenological) interview.
% todo add some more framing here to make it easier to step through the methods, maybe: For this paper we designed a prototype that lies between the two edge cases -- no intention and control to full control -- and investigate user's sense of agency using both explicit and implicit measures.

\subsection{Participants}
Nine participants (M = 30.00 years, SD = 3.81) were recruited from our local institution and through the institute's online participant pool. All participants were right-handed. Participants were compensated with course credit or 12 Euro per hour of study participation. Prior to their participation, they were informed of the nature of the experiment, recording, and anonymization procedures and signed a consent form. The experiment was approved by the ethics committee of [anonymized]. One participant had to be excluded from further data analyses due to significant deviations from the instructions in the execution of the task.
%the Department of Psychology and Ergonomics at TU Berlin (tracking number: BPN\_GEH\_2\_230130220421).

\subsection{Experimental Task, Design and Procedure}
Participants performed a simple tapping task in three conditions with 75 trials each: INTENTION, EXTERNAL, and AUGMENTED.
% , see figure~\ref{fig:task_design}.

@@

\indent\textbf{INTENTION.} The task went as follows: (1) a fixation cross appeared on a tablet screen and participants were instructed to rest and wait until it disappeared; (2) then, they were instructed to wait for a brief moment (2 to 3s), before (3) initiating their movement and tap the screen. In line with the literature on the origin of the RP generating process they were told ``to avoid pre-planning the movement, avoid any obvious rhythm over trials, and to press when they felt the spontaneous urge to move''~\cite{Schultze-Kraft2021-cu}. (4) After the screen was tapped, a tone was played at a pseudo-random delay of 200, 350, or 500ms. Participants were now asked to estimate the delay, typing in their answer on a number pad of an attached keyboard. After confirming their answer by hitting the return key, the next trial started.
During INTENTION, participants were equipped with Electromyogram (EMG) sensors instead of EMS electrodes. INTENTION data was used as training data for the BCI, see section~\ref{BCI} below, as well as to select stereotypical reaction times in the EXTERNAL condition.

@@

The order of the conditions was not pseudo-randomized since training data obtained in INTENTION was required for both EXTERNAL and AUGMENTED. Furthermore, AUGMENTED was always the last condition, thereby allowing for a prolonged interview.

% \begin{figure*}
%     \centering
%     \includegraphics[width=\textwidth]{figures/task_design.png}
%     \caption{(a) description of the experimental conditions, (b) their hypothesized effects on sense of agency from the perspective of a comparator model, is inspired by~\citet{Dewey2019-dg} (c) depicts the flow of movement generation in the participant and the system between conditions.}
%     \label{fig:task_design}
% \end{figure*}

\subsection{Apparatus}
The experimental setup, depicted in Figure~\ref{fig:setup}, comprised: (1) a 1-channel EMG device, (2) a 64-channel EEG system, (3) a medically-compliant EMS device connected via two electrodes worn on the forearm, and (4) a tablet to run the experiment and collect behavioral responses. To assist readers in replicating our experiment, we provide the necessary technical details, the complete source code for the experiment, the collected data, and the analysis scripts\footnote{[anonymized]}.
% TODO add links after anonymization is lifted.

\indent\textbf{(1) EMG Recording.} EMG data was recorded from 1 bipolar channel using the BrainAmp ExG amplifier (BrainProducts GmbH, Gilching, Germany). The two electrodes were placed above the flexor digitorum profundus with a reference electrode located on the wrist bone. EMG data was collected in synchrony with the EEG data through BrainProducts BrainVision Recorder.

\indent\textbf{(2) EEG Recording.} EEG data was recorded from 64 actively amplified Ag/AgCl electrodes in an actiCap Snap cap using BrainAmp DC amplifiers from BrainProducts. Electrodes were placed according to the extended international 10–20 system \cite{Jasper1983-uw}. One electrode was placed under the right eye to provide additional information about eye movements (vEOG). After fitting the cap, all electrodes were filled with conductive gel to ensure proper conductivity, and electrode impedance was brought below 10$\Omega$. EEG (and EMG) data were recorded with a sampling rate of 250 Hz.

\begin{figure}[!h]
    \centering
    \includegraphics[width=\columnwidth]{figures/setup.png}
    \caption{Experimental setup of measurement-- and input devices (image with consent from participant).}
    \label{fig:setup}
\end{figure}

\indent\textbf{(3) Electrical Muscle Stimulation.} We actuated the ring finger via EMS, which was delivered with two electrodes attached to the participants' flexor digitorum profundus muscle. We utilized the flexor digitorum profundus since we found that we can robustly actuate it without inducing unintended motion of neighboring fingers. This finger actuation was achieved via a medically-compliant battery powered muscle stimulator (TENS/EMS Super Duo Plus, prorelax, Düren, Germany). The EMS system's output was controlled by flipping a solid state relay (silent) connected via an Arduino Uno (Arduino, Monza, Italy) to the experiment computer. The EMS was pre-calibrated by the participant to ensure a pain-free but effective stimulation and robust actuation leading to an `immediate' tap on the touchscreen after actuation. To ensure a comfortable experimental experience so that participants relaxed their arm musculature as much as possible, a custom built hand rest (support device) was placed on top of the touchscreen for participants, see figure~\ref{fig:setup}.
% [todo] What were the settings of the stimulation?

\indent\textbf{(4) Experiment Presentation and Collection of Behavioral Responses.} An Acer Group (Acer Inc, Taipeh, Taiwan) tablet was used to present the task to participants and record their behavioral responses. In addition to the tablet, we used an external keyboard to allow response input of users indicating their timing judgements.

% alternative headings:
% Brain-Computer Interface to Detect EEG Readiness Potentials }
% Classifiying EEG Signals of the Intent to Interact
\subsection{Brain-Computer Interface}\label{BCI}
The data obtained in INTENTION was used to train the BCI. We utilized the [anonymized] and the EEGLAB~\cite{Delorme2004-sn} toolbox inside the MATLAB (The MathWorks Inc. Natick, MA, USA) environment for preprocessing both EEG and EMG data. First, to generate behavioral labels at a high temporal resolution for the EEG-based classifier, we leveraged EMG data from the flexor digitorum profundus. EMG amplitudes were band-pass filtered from 20 to 100 Hz and subsequently squared. Next, in order to label the time of movement onset, the EMG data was averaged across trials for the second preceding the screen tap. From this averaged data, the first sample where the EMG amplitude exceeded the 95th percentile was selected as the time of movement onset.
% bemobil pipeline

% TODO Next, noisy data segments were determined. 
% A segment was rejected as an outlier using Matlab's \textit{isoutlier} function using default parameters. % from isoutlier documentation (mean segment amplitude exceeding 3 scaled median absolute deviation)

@@

\subsubsection{Preprocessing EEG and Selecting discriminative channels}\label{eeg_methods}
The EEG data was band pass filtered from 0.1 to 15 Hz. In the first step to prepare the EEG data for classification, noisy data segments were rejected. To this end, the EEGLAB function `autorej' was used, keeping default parameters. A trial was excluded if \textit{either} data from the \textit{idle} or \textit{pre-movement} class was rejected.

Then, to decrease the dimensionality of the EEG data, a selection of discriminative channels was conducted following a selection approach outlined by~\citep{Schultze-Kraft2021-cu}. First, for both \textit{pre-movement} and \textit{idle} the mean signal in the last 100 ms of each segment was subtracted from the mean signal in the first 100 ms of the segment. The resulting value thus indicated how much the signal had changed in the course of the 1 s segments. In line with the literature, there should be a change over time in the \textit{pre-movement} segment but not in the \textit{idle} segment. Then, all channels were sorted (1) in descending order by the signal difference in the \textit{pre-movement} segments and (2) in ascending order by the signal difference in the \textit{idle} segments. With respect to the ability of the system to discriminate between movement intention and the absence of movement intention, an ideal channel should show a strong change in signal over the course of a \textit{pre-movement} segment but no difference during an \textit{idle} segment. The ranks of the two criteria were joined by summation and the resulting order was saved for later use in training and real-time application of the classifier. Lastly, channels C3, C4 and Cz were moved to the top of the order for every participant as these are most frequently reported in studies on the RP.

\subsubsection{Training of EEG classifier}
To classify EEG data a linear discriminant analysis (LDA) with shrinkage regularization (automatic shrinkage using the Ledoit-Wolf lemma~\cite{Ledoit2004-bi}) was trained for each participant individually. As single-trial features, the (linear) slope coefficient was obtained for both \textit{idle} and \textit{pre-movement} segments by fitting a linear regression. The classifier was then trained using the slope coefficient feature of the selection of the most discriminative channels as features, hence in the end, generating a feature vector in the dimension of channels that were kept for classification. We purposefully constrained the dimensionality of the feature vector to avoid over-fitting.

Using scikit-learn~\cite{Pedregosa2012-sj}, an LDA with automatic shrinkage was cross-validated (using 5-folds) for a grid search from 6 to 20 channels with a step size of adding 2 channels. Following this grid search, the cross-validation with the highest accuracy determined the number of channels that were kept for training the model for real-time application. Ultimately, to determine the threshold at which, during real-time application, the classifier would trigger the EMS, we computed the Receiver-operator characteristic (ROC) and from this selected the threshold at 15 \% false positive rate.

\subsubsection{Real-time application and EMS control}
During real-time application, the EEG data was buffered for the last second for the selected channels. The data was band-pass filtered analogously to the training data from 0.1 to 15 Hz. Next, the slope was computed for the discriminative channels selected in the classifier cross-validation grid search. This procedure ran at an update rate of 10 Hz, hence every 100 ms a new prediction was obtained from the classifier. To smooth the prediction output with the goal of reducing false predictions due to unlikely peaks, the predicted probability for the \textit{pre-movement} class was smoothed by averaging the last two predictions with a weighting (.3 and .5). Then, at 10 Hz update rate, this smoothed probability, as well as the predicted class for the current frame were gating the EMS switch: when the probability exceeded the threshold and the currently predicted class was \textit{pre-movement}, the switch was opened for 0.5 s.

@@

\subsection{Measures of Agency Experience: Intentional Binding, Question \& Interview}
To assess if SoA was preserved, an intentional binding measure, a standardized question, and an interview following the AUGMENTED block were investigated. At the conclusion of each condition we prompted participants to rate their experienced agency on a 7-point Likert scale with the statement ``It felt like I was in control of the movements during the task.'', the item was copied from~\cite{Bergstrom2022-fb}. Following AUGMENTED we interviewed users about their experience working with the system. After prompting users to recall their experience and summarize what their task had been we set the focus to the tapping movement and asked them to ignore the time estimation task for the following questions. We entered the open part of the interview by asking: ``What did the system do?'' followed up by ``What was the difference between the three conditions''. After some time, and depending on their answers, we reset the focus to the AUGMENTED condition and asked ``How often was the system active?'' followed by ``What do you think caused the actions of the system?''. This was then followed up by an `open' interview in which we frequently asked `how' and `why' questions to inquire about the user's experience.

% add back in after anonymization: All interviews were manually transcribed and then automatically translated using DeepL\footnote{https://www.deepl.com/translator}
We analyzed the interviews that always followed the AUGMENTED condition by loosely following~\citet{Mayring2015-pp}. All interviews were manually transcribed and [anonymized]. Prior to screening the texts, two experts clustered the responses into 3 clusters: First, ``Functionality'' referred to what participants attributed the source of the stimulation. Next, we clustered responses according to ``Guessed percentage'' of correct interaction, i.e., participants estimate of how well the system was aligned with their intention. For the cluster ``Correct Interaction'', we specifically queried participants to recall the moments where the stimulation felt in line with their intention to move, then we clustered their responses into positive and negative valence sentiments.

\subsection{Statistical Analyses}
To confirm and demonstrate the discriminative power of the EEG features, we plotted the amplitude time course of electrode Cz between \textit{pre-move} and \textit{idle} epochs. Next, the slope coefficients were extracted and a paired t-test was conducted. Ultimately, to assess the classifier performance, we report the F1 score, i.e. the harmonic mean between of the precision and recall, and plot the ROC.

\subsubsection{Hypotheses Testing}
Prior to any statistical analyses of the intentional binding measure, outlier trials were rejected. We applied `extreme outlier removal' using Tukey's method~\cite{Tukey1949-sl}. Three time intervals were considered relevant to describe `regular' behavior across trials which would manifest by: (1) Tapping the screen in a reasonable interval after the fixation cross disappeared. An excessively short or long delay indicated that participants either tapped the screen prematurely by accident or they were checking in with the experimenter, respectively. (2) Providing a `reasonable' estimation in the intentional binding task. (3) The EMS stimulation leading to an \textit{immediate} screen tap. A long delay between the EMS trigger and the subsequent screen tap indicated that the stimulation was not strong enough in this trial to lead to muscle actuation resulting in a screen tap. Taken together, applying Tukey's methods to these criteria led to the exclusion of 102 trials across all participants (M = 34.00, SD = 16.46) and a final data set of 1698 trials.

In line with the literature on intentional binding, we hypothesized that the time intervals should be underestimated for the INTENTION and the AUGMENTED conditions. This should not be the case for the EXTERNAL condition where users had no intention to move. Hence, when binding occurs, the intervals should be underestimated, hinting at a higher SoA. To test this, we fit a linear mixed effects model with \textit{condition} (INTENTION, EXTERNAL, AUGMENTED) as a fixed effect. As a random effect, we considered the intercepts for participants. We obtained p-values of the coefficients by calculating likelihood-ratio tests. All parameters were estimated by maximum likelihood estimation~\cite[see][]{Pinheiro2006-bk}. Post-hoc, pairwise contrasts were computed using estimated marginal means with Tukey's methods for multiple comparisons corrections~\cite{Lenth2020-xk}.
% ~\cite{winterLinearModelsLinear2013}

Next, we hypothesized that subjective ratings of control over the tap movement are comparable between INTENTION and AUGMENTED and lower in the EXTERNAL condition. Again, we fit a linear mixed effects model with \textit{condition} as a fixed effect and participant ID as a random effect. Coefficients were assessed in the same way as for the intentional binding parameter above.

In short, we tested two main hypotheses with regard to experiencing agency in our three experimental conditions: (1) participants underestimate the tone delay when acting intentionally. Adding EMS in line with participants' intention to move does not affect this underestimation. (2) The subjective feeling of control is comparable between INTENTION and AUGMENTED conditions. EXTERNAL should decrease the feeling of control significantly. Additionally, we report the clustered interviews anecdotally.

%%% Resources
% An additional cluster ``takeaways'' was used to cluster other, miscellaneous,  interesting comments.

>>>>

\section{User study \& Methods}
With this study, we wanted to find out whether the experience of agency can be preserved during physical action augmentation when the user's own brain signals serve as the control signal to a physical end effector. Therefore, we conducted a user study comparing three conditions of agency experience. With the first two conditions, INTENTION and EXTERNAL we queried users at the two edges of agency experience. In EXTERNAL participants had no intention to move and no control over their movement. On the other hand, in INTENTION, participants were acting as they would in their day-to-day lives, fully in control. We then investigated in a third condition where our AUGMENTED prototype sits on the spectrum between INTENTION and EXTERNAL, in other words, whether it preserved agency and to what degree. To answer our question, we employed a mixed methods approach including a psychometric test of intentional binding, one standardized question, and a qualitative (phenomenological) interview.

\subsection{Participants}
Nine participants (M = 30.00 years, SD = 3.81) were recruited from our local institution and through the institute's online participant pool. All participants were right-handed. Participants were compensated with course credit or 12 Euro per hour of study participation. Prior to their participation, they were informed of the nature of the experiment, recording, and anonymization procedures and signed a consent form. The experiment was approved by the ethics committee of [anonymized]. One participant had to be excluded from further data analyses due to significant deviations from the instructions in the execution of the task. Precisely, they did not initiate their movements at their own volition but rather immediately at the onset of each trial, thereby violating the task instruction of waiting 2-3 s after trial onset, see task description below for more detail.

\subsection{Apparatus}
The experimental setup, depicted in Figure~\ref{fig:setup}, comprised: (1) a 1-channel EMG device, (2) a 64-channel EEG system, (3) a medically-compliant EMS device connected to two electrodes worn on the forearm, and (4) a tablet to run the experiment and collect behavioral responses. To assist readers in replicating our experiment, we provide the necessary technical details, the complete source code for the experiment, the collected data, and the analysis scripts\footnote{[anonymized]}.
% TODO add links after anonymization is lifted.

\begin{figure}[!h]
    \centering
    \includegraphics[width=\columnwidth]{figures/setup_fixed.png}
    \caption{Experimental setup of measurement-- and input devices (image with consent from participant).}
    \label{fig:setup}
\end{figure}

\indent\textbf{(1) EMG Recording.} EMG data was recorded from 1 bipolar channel using the BrainAmp ExG amplifier (BrainProducts GmbH, Gilching, Germany). The two electrodes were placed above the flexor digitorum profundus with a reference electrode located on the wrist bone. EMG data was collected in synchrony with the EEG data through BrainProducts' BrainVision Recorder. EMG data was only recorded for the first experimental condition to obtain labels for classifier training. After that, the EMG electrodes were changed for EMS electrodes.

\indent\textbf{(2) EEG Recording.} EEG data was recorded from 64 actively amplified Ag/AgCl electrodes referenced to electrode FCz in an actiCap Snap cap using BrainAmp DC amplifiers from BrainProducts. Electrodes were placed according to the extended international 10–20 system \cite{Jasper1983-uw}. One electrode was placed under the right eye to provide additional information about eye movements (vEOG). After fitting the cap, all electrodes were filled with conductive gel to ensure proper conductivity, and electrode impedance was brought below 10k$\Omega$. EEG (and EMG) data were recorded with a sampling rate of 250 Hz.

\indent\textbf{(3) Electrical Muscle Stimulation.} We actuated the ring finger via EMS, which was delivered with two electrodes attached to the participants' flexor digitorum profundus muscle. We utilized the flexor digitorum profundus since we found that we can robustly actuate it without inducing unintended motion of neighboring fingers. This finger actuation was achieved via a medically compliant battery-powered muscle stimulator (TENS/EMS Super Duo Plus, prorelax, Düren, Germany). The EMS system's output was controlled by flipping a solid state relay (silent) connected via an Arduino Uno (Arduino, Monza, Italy) to the experiment computer. The EMS was pre-calibrated by the participant to ensure a pain-free but effective stimulation and robust actuation leading to an `immediate' tap on the touchscreen after actuation. To ensure a comfortable experimental experience so that participants relaxed their arm musculature as much as possible, a custom built hand rest (support device) was placed on top of the touchscreen for participants, see figure~\ref{fig:setup}.

\indent\textbf{(4) Experiment Presentation and Collection of Behavioral Responses.} An Acer Group (Acer Inc, Taipeh, Taiwan) tablet was used to present the task to participants and record their behavioral responses. In addition to the tablet, we used an external keyboard to allow users to input their timing judgments.

\subsection{Experimental Task, Design and Procedure}
Participants performed a simple tapping task in three conditions with 75 trials each: INTENTION, EXTERNAL, and AUGMENTED.

@@

\indent\textbf{INTENTION.} The task was as follows: (1) a fixation cross appeared on a tablet screen and participants were instructed to rest and wait until it disappeared; (2) they were instructed to wait for a brief moment (2 to 3s), before (3) initiating their movement and tap the screen, see figure~\ref{fig:progression}. In line with the literature on the origin of the RP generating process, they were told ``to avoid pre-planning the movement, avoid any obvious rhythm across trials, and to press when they felt the spontaneous urge to move''~\cite{Schultze-Kraft2021-cu}. (4) After the screen was tapped, a tone was played at a pseudo-random delay of 200, 350, or 500ms. Participants were now asked to estimate the delay, typing in their answers on a number pad of an attached keyboard. After confirming their answer by hitting the return key, the next trial started.

During INTENTION, participants were equipped with EMG sensors instead of EMS electrodes. Participants' reaction times in INTENTION were used to select stereotypical reaction times for the EXTERNAL condition. At the end of INTENTION, EMG electrodes were exchanged for EMS electrodes at the identical location on the forearm.

@@

The order of the conditions was not pseudo-randomized since training data obtained in INTENTION was required for both EXTERNAL and AUGMENTED. Furthermore, AUGMENTED was always the last condition, thereby allowing for a prolonged interview immediately after the experience with the prototype.

\subsection{Brain-Computer Interface}\label{BCI}
The data obtained in INTENTION was used to train the BCI. We utilized the [anonymized] and the EEGLAB~\cite{Delorme2004-sn} toolbox inside the MATLAB (The MathWorks Inc. Natick, MA, USA) environment for preprocessing both EEG and EMG data. First, to generate behavioral labels at a high temporal resolution for the EEG-based classifier, we leveraged EMG data from the flexor digitorum profundus. EMG amplitudes were band-pass filtered from 20 to 100 Hz and subsequently squared. Next, to label the time of movement onset, the EMG data was averaged across trials for the second preceding the screen tap. From this averaged data, the first sample where the EMG amplitude exceeded the 95th percentile was selected as the time of movement onset, see~\ref{fig:method} a.
% TODO remove anonymization bemobil pipeline

% TODO/Ignore Next, noisy data segments were determined. 
% A segment was rejected as an outlier using Matlab's \textit{isoutlier} function using default parameters. % from isoutlier documentation (mean segment amplitude exceeding 3 scaled median absolute deviation)

@@

\subsubsection{Preprocessing EEG and Selecting discriminative channels}\label{eeg_methods}
The EEG data was band pass filtered from 0.1 to 15 Hz. In the first step to prepare the EEG data for classifier training, noisy data segments were rejected. To this end, the EEGLAB function `autorej' was used, keeping default parameters. A trial was excluded if \textit{either} data from the \textit{idle} or \textit{pre-movement} class was rejected.

Then, to decrease the dimensionality of the EEG data, we selected discriminative channels following an approach outlined by~\citep{Schultze-Kraft2021-cu}. First, for both \textit{pre-movement} and \textit{idle} the mean signal in the last 100 ms of each segment was subtracted from the mean signal in the first 100 ms of the segment. The resulting value thus indicated how much the signal had changed in the course of the 1 s segments. In line with the literature, there should be a change over time in the \textit{pre-movement} segment but not in the \textit{idle} segment. Then, all channels were sorted (1) in descending order by the signal difference in the \textit{pre-movement} segments and (2) in ascending order by the signal difference in the \textit{idle} segments. With respect to the ability of the system to discriminate between movement intention and the absence of movement intention, an ideal channel should show a strong change in signal over the course of a \textit{pre-movement} segment but no difference during an \textit{idle} segment. The ranks of the two criteria were joined by summation and the resulting order was saved for later use in training and real-time application of the classifier. Lastly, channels C3, C4, and Cz were moved to the top of the order for every participant as these are most frequently reported in studies on the RP.

\begin{figure*}[!h]
    \centering
    \includegraphics[width=\textwidth]{figures/method.png}
    \caption{Diagram summarizing: (a) signal processing and event labeling from muscle activity, (b) extraction of linear slope features from labeled brain signals and (c) real-time application of the BCI gating the muscle stimulation hardware.}
    \label{fig:method}
\end{figure*}

\subsubsection{Training of EEG classifier}
To classify EEG data, a linear discriminant analysis (LDA) with shrinkage regularization (automatic shrinkage using the Ledoit-Wolf lemma~\cite{Ledoit2004-bi}) was trained for each participant individually. As single-trial features, the (linear) slope coefficient was obtained for both \textit{idle} and \textit{pre-movement} segments by fitting a linear regression. The classifier was then trained using the slope coefficient of the selection of the most discriminative channels as features, generating a feature vector in the dimension of channels that were kept for classification, see~\ref{fig:method} b. We purposefully constrained the dimensionality of the feature vector to avoid over-fitting and decrease the computational load.

Using scikit-learn~\cite{Pedregosa2012-sj}, the LDA was cross-validated (using 5-folds) for a grid search from 6 to 20 channels with a step size of adding 2 channels. Following this grid search, the cross-validation with the highest accuracy determined the number of channels that were kept for training the model for real-time application. Ultimately, to determine the threshold at which, during the real-time application, the classifier would trigger the EMS, we computed the Receiver-operator characteristic (ROC) and from this selected the threshold at 15 \% false positive rate.

\subsubsection{Real-time application and EMS control}
During real-time application, the EEG data was buffered for the last second for the selected discriminative channels. The data was band-pass filtered analogously to the training data from 0.1 to 15 Hz. Next, the slope was computed. This procedure ran at an update rate of 10 Hz, hence every 100 ms a new prediction was obtained from the classifier. To smooth the prediction output to reduce false predictions due to unlikely peaks, the predicted probability for the \textit{pre-movement} class was smoothed by averaging current and the preceding prediction with a weighting (weight of .3 for the preceding, and .5 for the current prediction). The weights were obtained through trial-and-error during piloting. Then, with a 10 Hz update rate, this smoothed probability, as well as the predicted class for the current frame were gating the EMS switch: when the probability exceeded the threshold and the currently predicted class was \textit{pre-movement}, the switch was opened for 0.5 s, see~\ref{fig:method} c.

@@

\subsection{Measures of Agency Experience: Intentional Binding, Question \& Interview}
To assess if SoA was preserved, an intentional binding measure, a standardized question, and an interview following the AUGMENTED block were investigated. After each condition we prompted participants to rate their experienced agency on a 7-point Likert scale with the statement ``It felt like I was in control of the movements during the task.'', the item was copied from~\cite{Bergstrom2022-fb}. Following AUGMENTED we interviewed users about their experience working with the system. After prompting users to recall their experience and summarize what their task had been, we set the focus to the tapping movement and asked them to ignore the time estimation task in the questions that followed. We entered the open part of the interview by asking: ``What did the system do?'' followed up by ``What was the difference between the three conditions''. After some time, and depending on their answers, we reset the focus to the AUGMENTED condition and asked ``How often was the system active?'' followed by ``What do you think caused the actions of the system?''. This was then followed up by an `open' interview in which we frequently asked `how' and `why' questions to inquire about the user's experience.

% add back in after anonymization: All interviews were manually transcribed and then automatically translated using }
We analyzed the interviews by loosely following~\citet{Mayring2015-pp}. All interviews were manually transcribed and translated to English using DeepL\footnote{https://www.deepl.com/translator} (DeepL SE, Cologne, Germany). Before screening the texts, two experts clustered the responses into 3 clusters: First, ``Functionality'' referred to what participants attributed the source of the stimulation. Next, we clustered responses according to ``Guessed percentage'' of correct interaction, i.e., participants' estimate of how well the system was aligned with their intention. For the cluster ``Correct Interaction'', we specifically queried participants to recall the moments where the stimulation felt in line with their intention to move, then we clustered their responses into sentiments with positive and negative valence.

\subsection{Statistical Analyses}
To confirm and demonstrate the discriminative power of the EEG features, we plotted the amplitude time course of electrode Cz between \textit{pre-move} and \textit{idle} epochs. Electrode Cz was chosen for exemplary presentation, as it is frequently reported in studies on the RP, since it captures neural activity originating from the sensorimotor cortex. Next, the slope coefficients were extracted and a paired t-test was conducted. Ultimately, to assess the classifier performance, we calculated the F1 score, i.e. the harmonic mean of precision and recall, and plotted the ROC, see figure~\ref{fig:EEG_results} a to c.

\subsubsection{Hypotheses Testing}
Prior to any statistical analyses of the intentional binding measure, outlier trials were rejected. We applied `extreme outlier removal' using Tukey's method~\cite{Tukey1949-sl}. Three time intervals were considered relevant to describe `regular' behavior across trials which would manifest by: (1) Tapping the screen in a reasonable interval after the fixation cross disappeared. An excessively short or long delay indicated that participants either tapped the screen prematurely by accident or they were checking in with the experimenter, respectively. (2) Providing a `reasonable' estimation in the intentional binding task. (3) The EMS stimulation leading to an \textit{immediate} screen tap. A long delay between the EMS trigger and the subsequent screen tap indicated that the stimulation was not strong enough in this trial to lead to muscle actuation resulting in a screen tap. Taken together, applying Tukey's method to each of these time windows and fusing the rejected trials led to the exclusion of 102 trials across all participants (M = 34.00, SD = 16.46) and a final data set of 1698 trials.

In line with the literature on intentional binding, we hypothesized that the time intervals should be underestimated for the INTENTION and the AUGMENTED conditions. This should not be the case for the EXTERNAL condition where users had no intention to move. Hence, when binding occurs, the intervals should be underestimated, hinting at a higher SoA. To test this, we fitted a linear mixed effects model with \textit{condition} (INTENTION, EXTERNAL, AUGMENTED) as a fixed effect and as a random effect, we considered the intercepts for participants ($time interval ~ condition + 1|participantID$). We obtained p-values of the coefficients by calculating likelihood-ratio tests. All parameters were estimated by maximum likelihood estimation~\cite{Pinheiro2006-bk}. Post hoc, we computed pairwise contrasts with the Tukey HSD (Honest Significant Differences), accounting for multiple companions with alpha correction~\cite{Lenth2020-xk}.

Next, we hypothesized that subjective ratings of control over the tap movement are comparable between INTENTION and AUGMENTED and lower in the EXTERNAL condition. Again, we fitted a linear mixed effects model with \textit{condition} as a fixed effect and participant ID as a random effect. Coefficients were assessed in the same way as for the intentional binding parameter above.

In short, we tested two main hypotheses with regard to the agency experience in our three experimental conditions: (1) participants underestimate the tone delay when acting intentionally. Adding EMS in line with participants' intention to move does not affect this underestimation. (2) The subjective feeling of control is comparable between INTENTION and AUGMENTED conditions. EXTERNAL should decrease the feeling of control significantly. Additionally, we report the clustered interviews anecdotally.

%%% Resources
\begin{comment}

% An additional cluster ``takeaways'' was used to cluster other, miscellaneous,  interesting comments.

### 2023-12-13 input/6_results.tex
revision

The grid-search over channels resulted in the BCI leveraging on average 11.5 (SD = 3.7) channels. \textcolor{green}{Besides channels C3, C4, and Cz, that were always included, other common channels (retained for at least 3 participants) included FT9, AF3, F5, F7, F2, FT8, and AF7.} The classifier cross-validation resulted in a mean F1 score of .7 (SD = .03), see figure~\ref{fig:EEG_results} c. We set the detection threshold to 15 \% false positive rate and at that rate, observed a mean threshold of 57 \% (SD = .02). Hence, on average, the classifier switched on the EMS when it predicted class \textit{pre-movement} with 57 \% probability.
% TODO add percentage of trials the prototype was activated before touching the screen (SD = ?)

\subsection{Subjective Reports}
Six out of the eight participants made a reference to the source of the EMS in the AUGMENTED condition. In that subgroup, most participants wondered how the system worked, one remarked, ``I can't explain how it works technically.''. They found the stimulation to be sporadic, even during their own actions, leading to a belief that it was coincidental. As one participant put it, ``I have no idea what controlled the stimulation, I think it was coincidence when stimulation occurred.''. Furthermore, participants reported that they found the timing of the stimulation to be unpredictable, with one participant noting, ``Stimulation was random in time.''. Some perceived the stimulation as externally triggered, yet partially responsive to their choices, resulting in a sentiment captured by, ``I think that the stimulation in the third block was partly involuntary, but partly as if it was following my decision / choice.''. One participant briefly considered the involvement of their brain waves in controlling the stimulation, but expressing doubts about this possibility, stating, ``But I don't think that is the case.''.

>>>>

The grid-search over channels resulted in the BCI leveraging on average 11.5 (SD = 3.7) channels.Besides channels C3, C4, and Cz, that were always included, other common channels (retained for at least 3 participants) included FT9, AF3, F5, F7, F2, FT8, and AF7. The classifier cross-validation resulted in a mean F1 score of .7 (SD = .03), see figure~\ref{fig:EEG_results} c. We set the detection threshold to 15 \% false positive rate and at that rate, observed a mean threshold of 57 \% (SD = .02). Hence, on average, the classifier switched on the EMS when it predicted class \textit{pre-movement} with 57 \% probability.

\subsection{Subjective Reports}
Six out of the eight participants made a reference to the source of the EMS in the AUGMENTED condition. In that subgroup, most participants wondered how the system worked, one remarked, ``I can't explain how it works technically.''. They found the stimulation to be sporadic, even during their own actions, leading to a belief that it was coincidental. As one participant put it, ``I have no idea what controlled the stimulation, I think it was coincidence when stimulation occurred.''. Furthermore, participants reported that they found the timing of the stimulation to be unpredictable, with one participant noting, ``Stimulation was random in time.''. Some perceived the stimulation as externally triggered, yet partially responsive to their choices, resulting in a sentiment captured by, ``I think that the stimulation in the third block was partly involuntary, but partly as if it was following my decision/choice.''. One participant briefly considered the involvement of their brain waves in controlling the stimulation, but expressed doubts about this possibility, stating, ``But I don't think that is the case.''.
